---
title: "p8105_hw5_khs2318"
author: "Kayla Schiffer-Kane"
date: "2023-11-15"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(p8105.datasets)
library(tidyverse)
library(ggridges)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1 

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

## Load Data

**Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).**

```{r load_data}
data = read_csv('./data/homicide-data.csv') |>
  janitor::clean_names() |>
  mutate(city_state = paste(city,state, sep = ", "))

```


The data has `r nrow(data)` rows and `r ncol(data)` columns. The data includes character variables with details on the victims, including their names, race, age, and sex. It also includes variables for the city and state where the murder happened. There are `r data |> pull(city) |> unique() |> length()` cities and `r data |> pull(state) |> unique() |> length()` states. The data also includes disposition, which describes the status of the case, and details on reporting including date, latitude, and longitude. 


```{r summarize}
data_prop = data |> 
  group_by(city, city_state) |>
  summarize(n_cases = n(), 
            n_unsolved = 
              sum(disposition %in% c("Closed without arrest","Open/No arrest")))
data_prop |>
  knitr::kable()
```


## Proportion of Unsolved Homicides

### Baltimore
**For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of `prop.test` as an R object, apply the `broom::tidy` to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.**

```{r prop_bt}
data_prop_bt = data_prop |>
  filter(city_state == "Baltimore, MD")

prop.test(data_prop_bt$n_unsolved, data_prop_bt$n_cases) |>
  broom::tidy() |>
  select(estimate, conf.low, conf.high)

data_prop |>
  filter(city_state == "Baltimore, MD") |>
  mutate(prop = map2(n_unsolved, n_cases, ~broom::tidy(prop.test(.x,.y)))) |>
  unnest(prop) |>
  select(estimate,conf.low,conf.high)
```


### All Cities

**Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of `purrr::map`, `purrr::map2`, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.**

```{r prop_all}
city_prop = data_prop |>
  mutate(prop = map2(n_unsolved, n_cases, ~broom::tidy(prop.test(.x,.y)))) |>
  unnest(prop) |>
  select(city, city_state, n_cases, n_unsolved, estimate,conf.low,conf.high)
city_prop
```


## Plot CIs by city

Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r city_plot}
city_prop |>
  ggplot(aes(x = reorder(city, estimate), y = estimate)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```




